Associations between Chlorophyll \emph{a} and various Microcystin-LR Health Advisory Concentrations

### 2015-10-16 manuscript/manuscript.tex
Betty's edits

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Changes borrowed from @cboettig, added by @jhollist 
% A modified page layout 
\textwidth 6.75in
\oddsidemargin -0.15in
\evensidemargin -0.15in
\textheight 9in
\topmargin -0.5in
\usepackage{lineno} % add 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Cyanobacteria harmful algal blooms (cHABs) are associated with a wide
range of adverse health effects that stem mostly from the presence of
cyanotoxins. To help protect against these impacts, several health
advisory levels have been set for some toxins. In particular, one of the
more common toxins, microcystin, has several advisory levels set for
drinking water and recreational use. However, compared to other water
quality measures, field measurements of microcystin are not commonly
available due to cost and require advanced understanding to interpret
results. Addressing these issues will take time and resources. Thus,
there is utility in finding indicators of microcystin that are already
widely available, can be estimated quickly and \emph{in situ}, and used
as a first defense against high levels of microcystin. In particular,
chlorophyll \emph{a} is very commonly measured, can be estimated
\emph{in situ}, and has been shown to be positively associated with
microcystin. In this paper we use this association to provide estimates
of chlorophyll \emph{a} that if exceeded would be indicative of a higher
probability of exceeding select health advisory concentrations for
microcystin-LR. Using the 2007 National Lakes Assessment and a
conditional probability approach that has been used in other water
quality settings, we identify chlorophyll \emph{a} concentrations that
are more likely than not to be associated with an exceedance of a
microcystin health advisory level. We look at the recent US EPA
standards for drinking water as well as the World Health Organization
levels for drinking water and recreational use. For microcystin
concentrations of 0.3, 1, 1.6, and 2 we find chlorophyll \emph{a}
concentrations of 22.25, 65.6, 84.96, and 105.84, respectively. When
managing for these various microcystin levels exceeding these reported
chlorophyll \emph{a} concentrations should be a trigger for further
testing and possible management action.

@@

Over the last decade or so numerous events and legislative activities
have raised the public awareness of harmful algal blooms (Rinta-Kanto et
al. 2009, HABHRCA 2014, Jetoo et al. 2015), and the US Environmental
Protection Agency (USEPA) has recently released suggested microcystin-LR
(one of the more common toxins) concentrations that would trigger health
advisories (McElhiney and Lawton 2005, Zurawell et al. 2005, USEPA
2015). Additionally, the World Health Organization (WHO) has had
proposed advisory levels for drinking water and a range of recreational
risk levels (Chorus and Bartram 1999, (WHO) and others 2003). While
these levels and associated advisories are likely to help mitigate the
impacts from harmful algal blooms, they are not without complications.

One of these complications is that they rely on available measurements
of microcystin-LR. While laboratory testing remains the gold standard
for quantifying microcystin-LR concentrations in water samples, several
field test kits have been developed. Even though field tests provide a
much needed means for rapid assessment, they are not yet widely used and
are moderately expensive (approximately \$150-\$200 depending on
specific kit) with a limited shelf life (typically one year) (James et
al. 2011, Aranda-Rodriguez et al. 2015). Additionally, each technique
requires nuanced understanding of the detection method (e.g., limit of
detection, specific microcystin variants being measured, and sampling
protocol). Fortunately, microcystin-LR has been shown to be associated
with several other, more commonly measured and understood components of
water quality.

Chlorophyll \emph{a} is a very commonly measured component of water
quality that is also known to be associated with Microsystin-LR
concentrations {[}Lee et al. (2000); Paerl and Otten (2013);
heisler2008eutrophication{]}. Additionally there are many rapid
measurements for assessing chlorophyll \emph{a} levels \emph{in situ}.
For instance, there are small or hand held flourometers that measure
chlorohpyll \emph{a}. Given these facts, it might be possible to
identify chlorophyll \emph{a} concentrations that would be associated
with the various Microcystin-LR health advisory levels. Identifying
these associations would provide another tool for water resource
managers to use to help manage the threat to public health posed by
cHABs and would be especially useful in the absence of microcystin-LR
concentrations. Thus, the goal of this paper is to utilize the National
Lakes Assessment data and identify chlorophyll \emph{a} concentrations
that are associated with higher probabilities of exceeding several
microcystin-LR health advisory concentrations (USEPA 2009). So that
others may repeat or adjust this analysis, the data, code, and this
manuscript are freely available via
\href{https://github.com/USAPE/microcystinchla}{\url{https://github.com/USAPE/microcystinchla}}.

@@

We used the 2007 National Lakes Assessment (NLA) water quality and
microcystin-LR concentration data (USEPA 2009). These data represent a
snapshot of water quality from the summer of 2007 and data on
chlorophyll \emph{a} and microcystin-LR concentrations are available for
lakes.

@@

We used a conditional probability analysis (CPA) approach to explore
associations between chlorophyll \emph{a} concentrations and World
Health Organization (WHO) and U.S. Environmental Protection Agency (U.S.
EPA) microcystin-LR health advisory levels (Paul and Munns 2011). Many
levels have been suggested (Table \ref{tab:microcystin_levels}), but
lakes with higher microcystin-LR concentrations in the NLA were rare.
Only 1.16 \% of lakes sampled had a concentration greater than 10. Thus,
for this analysis we focus on the microcystin concentrations that are
better represented in the NLA data. These were 0.3, 1, 1.6, and 2
\(\mu\)g/L.

A detailed discussion of CPA is beyond the scope of this paper, but see
Paul et al. (2005) and Hollister et al. (2008) for details. For this
analysis, we used CPA to examine how the conditional probability of
exceeding one of the health advisory changes as chlorophyll \emph{a}
increases in a lake. The 95\% confidence intervals were calculated from
1000 bootstrapped samples. To identify chlorophyll \emph{a}
concentrations of concern we used a 50\% conditional probability of
exceeding each health advisory level and extracted the minimum
chlorophyll \emph{a} concentration that was associated with an upper
confidence level of 50\% or greater. As both microcystin-LR and
chlorophyll \emph{a} values were highely skewed right, a log base 10
transformation was used. Additional details of the specific
implementation are available at
\url{https://github.com/USEPA/microcystinchla}.

@@

In the 2007 NLA, microcystin-LR concentrations ranged from 0.05 to 225.
Microcystin-LR concentrations of 0.05 \(\mu\)g/L represent the detection
limits. Any value greater than that indicates the presence of
microcystin-LR. Of those lakes with microcystin, the median
concentration was 0.51 and the mean was 3.1661047. Lastly, of all lakes
sampled, 21\% of lakes exceeded the U.S. EPA childrens drinking water
standard, 8.8\% of lakes exceeded the U.S. EPA adult drinking water
standard,11.7\% of lakes exceeded the WHO drinking water standard,and
7.3\% of lakes exceeded the WHO recreational standard. For chlorophyll
\emph{a}, the range was 0.07 to 936 \(\mu\)g/L. All lakes had reported
chlorophyll \emph{a} concentrations that exceeded detection limits. The
median concentration was 7.79 and the mean was 29.63. The associations
between chlorophyll \emph{a} and the upper confidence interval with a
conditional probability of 50\% were 22.25, 65.6, 84.96, and 105.84 for
0.3, 1.0, 1.6 and 2.0 \(\mu\)g/L, respectively (Table
\ref{tab:mc_chla_table} \& Figure \ref{fig:multi_cp_plot}).

@@

The association between Log 10 microcystin-LR and Log 10 chlorophyll
\emph{a} show a wedge pattern (Figure \ref{fig:chla_micro_scatter}).
This indicates that higher concentrations of microcystin-LR almost
always co-occur with higher concentrations of chlorophyll \emph{a} yet
the inverse is not true. Higher chlorophyll \emph{a} is not necessarily
predictive of higher microcystin-LR concentrations; however, chlorophyll
\emph{a} may be predictive of the probability of exceeding a certain
concentration.

This is the case as the probability of exceeding each of the four tested
health advisory levels increases as a function of chlorophyll \emph{a}
concentration (Figure \ref{fig:multi_cp_plot}). We use this association
to identify chlorophyll \emph{a} concentrations that are associated with
greater than even odds of exceeding a given health advisory level (Table
\ref{tab:mc_chla_table}). These represent 29.1\%, 11.5\%, 8.6\%, and
6.9\% of sample lakes for the U.S EPA childrens drinking water, the WHO
drinking water, the U.S. EPA Adult drinking water, and the WHO
recreational advisories, respecitvely.

Furthermore, the chlorophyll \emph{a} cutoffs may be used to predict
whether or not a lake exceeds the microcystin-LR health advisories.
Doing so allows us to compare the accuracy of the prediction as well as
evaluate false negatives. Total accuracy of the four cutoffs predicting
microcystin-LR exceedances were 74\% for the USEPA children's advisory,
86\% for the WHO drinking water advisory, 90\% for the USEPA adult
advisory and, 91\% for the WHO regreational advisory (Tables
\ref{tab:child_conmat_table}, \ref{tab:who_drink_conmat_table},
\ref{tab:adult_conmat_table}, \& \ref{tab:who_rec_conmat_table}). But
Total Accuracy is one part of the prediction performace with which we
are concerned.

When using the chlorophyll \emph{a} cutoffs as an indicator of
microcystin-LR exceedances, the error that should be avoided is
predicting that no exceedance has occurred when in fact it has. In other
words, we would like to avoid Type II errors and minimize the proportion
of false negatives. For the four chlorophyll \emph{a} cut-offs we had a
proportion of false negatives of 9\%, 7\%, 6\% and , 5\% for the U.S EPA
childrens drinking water, the WHO drinking water, the U.S. EPA Adult
drinking water, and the WHO recreational advisories, respecitvely. In
each case we miss less than 10\% of the lakes that are in fact exceeding
the microcystin-LR advisory. This is an acceptable rate, especially when
no other tests are available for a given water body.

@@

\begin{figure}[htbp]
\centering
\includegraphics{manuscript_files/figure-latex/epa_child_cp_plot-1.pdf}
\caption{Conditional probability plots showing association between the
probability of exceeding various microcystin-LR (MLR) health advisory
Levels. A.) Plot for U.S. EPA childrens drinking water advisory (0.3
\(\mu\)g/L). B.) Plot for WHO drinking water advisory (1 \(\mu\)g/L).
C.) Plot for U.S. EPA adult drinking water advisory (1.6 \(\mu\)g/L).
D.) Plot for WHO recreational advisory (2 \(\mu\)g/L).
\label{fig:multi_cp_plot}}
\end{figure}

@@

\begin{longtable}[c]{@{}lll@{}}
\caption{Various suggested microcystin-LR health advisory
concentrations. \label{tab:microcystin_levels}}\tabularnewline
\toprule
Source & Type & Concentration\tabularnewline
\midrule
\endfirsthead
\toprule
Source & Type & Concentration\tabularnewline
\midrule
\endhead
WHO & Drinking & 1 \(\mu\)g/L\tabularnewline
U.S. EPA & Drinking & 0.3 \(\mu\)g/L\tabularnewline
U.S. EPA & Drinking & 1.6 \(\mu\)g/L\tabularnewline
WHO & Recreational & 2-4 \(\mu\)g/L\tabularnewline
WHO & Recreational & 10-20 \(\mu\)g/L\tabularnewline
WHO & Recreational & 20-2000 \(\mu\)g/L\tabularnewline
WHO & Recreational & \textgreater{}2000 \(\mu\)g/L\tabularnewline
\bottomrule
\end{longtable}

@@

\begin{longtable}[c]{@{}llrr@{}}
\caption{Chlorophyll \textit{a} concentrations that are associated with
a 50\% probability of exceeding a microcystin-LR health advisory
concentration. \label{tab:mc_chla_table}}\tabularnewline
\toprule
Source & Type & Microcystin & Chlorophyll\tabularnewline
\midrule
\endfirsthead
\toprule
Source & Type & Microcystin & Chlorophyll\tabularnewline
\midrule
\endhead
U.S. EPA & Drinking & 0.3 & 22.25\tabularnewline
WHO & Drinking & 1.0 & 65.60\tabularnewline
U.S. EPA & Drinking & 1.6 & 84.96\tabularnewline
WHO & Recreational & 2.0 & 105.84\tabularnewline
\bottomrule
\end{longtable}

@@

\begin{longtable}[c]{@{}lrr@{}}
\caption{Confusion matrix comparing chlorophyll \textit{a} predicted
exceedences (rows) versus real exceedances (columns) for the U.S. EPA
childrens drinking water advisory.
\label{tab:child_conmat_table}}\tabularnewline
\toprule
& FALSE & TRUE\tabularnewline
\midrule
\endfirsthead
\toprule
& FALSE & TRUE\tabularnewline
\midrule
\endhead
FALSE & 638 & 94\tabularnewline
TRUE & 173 & 123\tabularnewline
\bottomrule
\end{longtable}

@@

\begin{longtable}[c]{@{}lrr@{}}
\caption{Confusion matrix comparing chlorophyll \textit{a} predicted
exceedences (rows) versus real exceedances (columns) for the WHO
drinking water advisory.
\label{tab:who_drink_conmat_table}}\tabularnewline
\toprule
& FALSE & TRUE\tabularnewline
\midrule
\endfirsthead
\toprule
& FALSE & TRUE\tabularnewline
\midrule
\endhead
FALSE & 838 & 76\tabularnewline
TRUE & 69 & 45\tabularnewline
\bottomrule
\end{longtable}

@@

\begin{longtable}[c]{@{}lrr@{}}
\caption{Confusion matrix comparing chlorophyll \textit{a} predicted
exceedences (rows) versus real exceedances (columns) for the U.S. EPA
adult drinking water advisory.
\label{tab:adult_conmat_table}}\tabularnewline
\toprule
& FALSE & TRUE\tabularnewline
\midrule
\endfirsthead
\toprule
& FALSE & TRUE\tabularnewline
\midrule
\endhead
FALSE & 887 & 57\tabularnewline
TRUE & 50 & 34\tabularnewline
\bottomrule
\end{longtable}

@@

\begin{longtable}[c]{@{}lrr@{}}
\caption{Confusion matrix comparing chlorophyll \textit{a} predicted
exceedences (rows) versus real exceedances (columns) for the WHO
recreational water advisory.
\label{tab:who_rec_conmat_table}}\tabularnewline
\toprule
& FALSE & TRUE\tabularnewline
\midrule
\endfirsthead
\toprule
& FALSE & TRUE\tabularnewline
\midrule
\endhead
FALSE & 910 & 52\tabularnewline
TRUE & 43 & 23\tabularnewline
\bottomrule
\end{longtable}

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Changes borrowed from @cboettig, added by @jhollist 
% A modified page layout 
\textwidth 6.75in
\oddsidemargin -0.15in
\evensidemargin -0.15in
\textheight 9in
\topmargin -0.5in
\usepackage{lineno} % add 
  \linenumbers % turns line numbering on 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Cyanobacteria harmful algal blooms (cHABs) are associated with a wide
range of adverse health effects that stem mostly from the presence of
cyanotoxins. To help protect against these impacts, several health
advisory levels have been set for some toxins. In particular, one of the
more common toxins, microcystin, has several advisory levels set for
drinking water and recreational use. However, compared to other water
quality measures, field measurements of microcystin are not commonly
available due to cost and advanced understanding required to interpret
results. Addressing these issues will take time and resources. Thus,
there is utility in finding indicators of microcystin that are already
widely available, can be estimated quickly and \emph{in situ}, and used
as a first defense against high levels of microcystin. In particular,
chlorophyll \emph{a} is very commonly measured, can be estimated
\emph{in situ}, and has been shown to be positively associated with
microcystin. In this paper, we use this association to provide estimates
of chlorophyll \emph{a} concentrations that if exceeded would be
indicative of a higher probability of exceeding select health advisory
concentrations for microcystin-LR. Using the 2007 National Lakes
Assessment and a conditional probability approach, we identify
chlorophyll \emph{a} concentrations that are more likely than not to be
associated with an exceedance of a microcystin health advisory level. We
look at the recent US EPA standards for drinking water as well as the
World Health Organization levels for drinking water and recreational
use. For the specific advisory microcystin concentrations of 0.3, 1,
1.6, and 2, we find chlorophyll \emph{a} concentrations of 23.68, 63.94,
84.96, and 97.49, respectively. When managing for these various
microcystin levels, exceeding these reported chlorophyll \emph{a}
concentrations should be a trigger for further testing and possible
management action.

@@

Over the last decade or so numerous events and legislative activities
have raised the public awareness of harmful algal blooms (Rinta-Kanto et
al. 2009, HABHRCA 2014, Jetoo et al. 2015), and in response the US
Environmental Protection Agency (USEPA) has recently released suggested
microcystin-LR (one of the more common toxins) concentrations that would
trigger health advisories (McElhiney and Lawton 2005, Zurawell et al.
2005, USEPA 2015). Additionally, the World Health Organization (WHO) has
had proposed advisory levels for drinking water and a range of
recreational risk levels (Chorus and Bartram 1999, (WHO) and others
2003). While these levels and associated advisories are likely to help
mitigate the impacts from harmful algal blooms, they are not without
complications.

One of these complications is that they rely on available measurements
of microcystin-LR. While laboratory testing remains the gold standard
for quantifying microcystin-LR concentrations in water samples, several
field test kits have been developed. Even though field tests provide a
much needed means for rapid assessment, they are not yet widely used and
are moderately expensive (approximately \$150-\$200 depending on
specific kit) with a limited shelf life (typically one year) (James et
al. 2011, Aranda-Rodriguez et al. 2015). Additionally, each technique
requires nuanced understanding of the detection method (e.g., limit of
detection, specific microcystin variants being measured, and sampling
protocol). Fortunately, microcystin-LR has been shown to be associated
with several other, more commonly measured and well understood
components of water quality.

Chlorophyll \emph{a} is a very commonly measured component of water
quality that is also known to be associated with Microsystin-LR
concentrations (Lee et al. 2000, Heisler et al. 2008, Paerl and Otten
2013). Additionally, there are many rapid measurements for assessing
chlorophyll \emph{a} levels \emph{in situ}. For instance, there are
small or hand held flourometers that measure chlorohpyll \emph{a}. Given
these facts, it might be possible to identify chlorophyll \emph{a}
concentrations that would be associated with the various Microcystin-LR
health advisory levels. Identifying these associations would provide
another tool for water resource managers to help manage the threat to
public health posed by cHABs and would be especially useful in the
absence of microcystin-LR concentrations. Thus, the goal of this paper
is to utilize the National Lakes Assessment (NLA) data and identify
chlorophyll \emph{a} concentrations that are associated with higher
probabilities of exceeding several microcystin-LR health advisory
concentrations (USEPA 2009). So that others may repeat or adjust this
analysis, the data, code, and this manuscript are freely available via
\href{https://github.com/USAPE/microcystinchla}{\url{https://github.com/USAPE/microcystinchla}}.

@@

We used the 2007 NLA water quality and microcystin-LR concentration data
(USEPA 2009). These data represent a snapshot of water quality from the
summer of 2007 and data on chlorophyll \emph{a} and microcystin-LR
concentrations are available for lakes.

@@

We used a conditional probability analysis (CPA) approach to explore
associations between chlorophyll \emph{a} concentrations and World
Health Organization (WHO) and USEPA microcystin-LR health advisory
levels (Paul and Munns 2011). Many levels have been suggested (Table
\ref{tab:microcystin_levels}), but lakes with higher microcystin-LR
concentrations in the NLA were rare. Only 1.16 \% of lakes sampled had a
concentration greater than 10. Thus, for this analysis we focus on the
microcystin concentrations that are better represented in the NLA data.
These were 0.3, 1, 1.6, and 2 \(\mu\)g/L.

A detailed discussion of CPA is beyond the scope of this paper; see Paul
et al. (2005) and Hollister et al. (2008) for greater detail. For this
analysis, we used CPA to examine how the conditional probability of
exceeding one of the health advisory changes as chlorophyll \emph{a}
increases in a lake. The 95\% confidence intervals were calculated from
1000 bootstrapped samples. To identify chlorophyll \emph{a}
concentrations of concern we used a 50\% conditional probability of
exceeding each health advisory level and extracted the minimum
chlorophyll \emph{a} concentration that was associated with an upper
confidence level of 50\% or greater. As both microcystin-LR and
chlorophyll \emph{a} values were highely skewed right, a log base 10
transformation was used. Additional details of the specific
implementation are available at
\url{https://github.com/USEPA/microcystinchla}.

@@

In the 2007 NLA, microcystin-LR concentrations ranged from 0.05 to 225.
Microcystin-LR concentrations of 0.05 \(\mu\)g/L represent the detection
limits. Any value greater than that indicates the presence of
microcystin-LR. Of those lakes with microcystin, the median
concentration was 0.51 and the mean was 3.17. Of all lakes sampled, 21\%
of lakes exceeded the USEPA childrens drinking water standard, 8.8\% of
lakes exceeded the USEPA adult drinking water standard,11.7\% of lakes
exceeded the WHO drinking water standard,and 7.3\% of lakes exceeded the
WHO recreational standard. For chlorophyll \emph{a}, the range was 0.07
to 936 \(\mu\)g/L. All lakes had reported chlorophyll \emph{a}
concentrations that exceeded detection limits. The median concentration
was 7.79 and the mean was 29.63. The associations between chlorophyll
\emph{a} and the upper confidence interval with a conditional
probability of 50\% were 23.68, 63.94, 84.96, and 97.49 for 0.3, 1.0,
1.6 and 2.0 \(\mu\)g/L advisory levels, respectively (Table
\ref{tab:mc_chla_table} \& Figure \ref{fig:multi_cp_plot}).

@@

The association between Log10 microcystin-LR and Log10 chlorophyll
\emph{a} shows a wedge pattern (Figure \ref{fig:chla_micro_scatter}).
This indicates that higher concentrations of microcystin-LR almost
always co-occur with higher concentrations of chlorophyll \emph{a} yet
the inverse is not true. Higher chlorophyll \emph{a} is not necessarily
predictive of higher microcystin-LR concentrations; however, chlorophyll
\emph{a} may be predictive of the probability of exceeding a certain
concentration.

This is the case as the probability of exceeding each of the four tested
health advisory levels increases as a function of chlorophyll \emph{a}
concentration (Figure \ref{fig:multi_cp_plot}). We use this association
to identify chlorophyll \emph{a} concentrations that are associated with
greater than even odds of exceeding a given health advisory level (Table
\ref{tab:mc_chla_table}). These represent 27.8\%, 11.9\%, 8.6\%, and
7.9\% of sample lakes for the U.S EPA childrens drinking water, the WHO
drinking water, the USEPA Adult drinking water, and the WHO recreational
advisories, respecitvely.

Furthermore, the chlorophyll \emph{a} cutoffs may be used to predict
whether or not a lake exceeds the microcystin-LR health advisories.
Doing so allows us to compare the accuracy of the prediction as well as
evaluate false negatives. Total accuracy of the four cutoffs predicting
microcystin-LR exceedances were 75\% for the USEPA children's advisory,
86\% for the WHO drinking water advisory, 90\% for the USEPA adult
advisory and, 90\% for the WHO regreational advisory (Tables
\ref{tab:child_conmat_table}, \ref{tab:who_drink_conmat_table},
\ref{tab:adult_conmat_table}, \& \ref{tab:who_rec_conmat_table}).
However, total accuracy is only one part of the prediction performace
with which we are concerned.

When using the chlorophyll \emph{a} cutoffs as an indicator of
microcystin-LR exceedances, the error that should be avoided is
predicting that no exceedance has occurred when in fact it has. In other
words, we would like to avoid Type II errors and minimize the proportion
of false negatives. For the four chlorophyll \emph{a} cut-offs we had a
proportion of false negatives of 9\%, 7\%, 6\% and , 5\% for the U.S EPA
childrens drinking water, the WHO drinking water, the USEPA adult
drinking water, and the WHO recreational advisories, respecitvely. In
each case we miss less than 10\% of the lakes that are in fact exceeding
the microcystin-LR advisory.

@@

\begin{figure}[htbp]
\centering
\includegraphics{manuscript_files/figure-latex/epa_child_cp_plot-1.pdf}
\caption{Conditional probability plots showing association between the
probability of exceeding various microcystin-LR (MLR) health advisory
Levels. A.) Plot for USEPA childrens drinking water advisory (0.3
\(\mu\)g/L). B.) Plot for WHO drinking water advisory (1 \(\mu\)g/L).
C.) Plot for USEPA adult drinking water advisory (1.6 \(\mu\)g/L). D.)
Plot for WHO recreational advisory (2 \(\mu\)g/L).
\label{fig:multi_cp_plot}}
\end{figure}

@@

\begin{longtable}[c]{@{}lll@{}}
\caption{Various suggested microcystin-LR health advisory
concentrations. \label{tab:microcystin_levels}}\tabularnewline
\toprule
Source & Type & Concentration\tabularnewline
\midrule
\endfirsthead
\toprule
Source & Type & Concentration\tabularnewline
\midrule
\endhead
USEPA & Adult Drinking Water Advisory & 1.6 \(\mu\)g/L\tabularnewline
USEPA & Child Drinking Water Advisory & 0.3 \(\mu\)g/L\tabularnewline
WHO & Drinking Water & 1 \(\mu\)g/L\tabularnewline
WHO & Recreational: High Prob. of Effect & 20-2000
\(\mu\)g/L\tabularnewline
WHO & Recreational: Low Prob. of Effect & 2-4 \(\mu\)g/L\tabularnewline
WHO & Recreational: Moderate Prob. of Effect & 10-20
\(\mu\)g/L\tabularnewline
WHO & Recreational: Very High Prob. of Effect & \textgreater{}2000
\(\mu\)g/L\tabularnewline
\bottomrule
\end{longtable}

@@

\begin{longtable}[c]{@{}llrr@{}}
\caption{Chlorophyll \textit{a} concentrations that are associated with
a 50\% probability of exceeding a microcystin-LR health advisory
concentration. \label{tab:mc_chla_table}}\tabularnewline
\toprule
Source & Type & Microcystin & Chlorophyll\tabularnewline
\midrule
\endfirsthead
\toprule
Source & Type & Microcystin & Chlorophyll\tabularnewline
\midrule
\endhead
USEPA & Child Drinking Water Advisory & 0.3 & 23.68\tabularnewline
WHO & Drinking Water Advisory & 1.0 & 63.94\tabularnewline
USEPA & Adult Drinking Water & 1.6 & 84.96\tabularnewline
WHO & Recreational: Low Prob. of Effect & 2.0 & 97.49\tabularnewline
\bottomrule
\end{longtable}

@@

\begin{longtable}[c]{@{}lrr@{}}
\caption{Confusion matrix comparing chlorophyll \textit{a} predicted
exceedences (rows) versus real exceedances (columns) for the USEPA
childrens drinking water advisory.
\label{tab:child_conmat_table}}\tabularnewline
\toprule
& FALSE & TRUE\tabularnewline
\midrule
\endfirsthead
\toprule
& FALSE & TRUE\tabularnewline
\midrule
\endhead
FALSE & 650 & 96\tabularnewline
TRUE & 161 & 121\tabularnewline
\bottomrule
\end{longtable}

@@

\begin{longtable}[c]{@{}lrr@{}}
\caption{Confusion matrix comparing chlorophyll \textit{a} predicted
exceedences (rows) versus real exceedances (columns) for the WHO
drinking water advisory.
\label{tab:who_drink_conmat_table}}\tabularnewline
\toprule
& FALSE & TRUE\tabularnewline
\midrule
\endfirsthead
\toprule
& FALSE & TRUE\tabularnewline
\midrule
\endhead
FALSE & 835 & 75\tabularnewline
TRUE & 72 & 46\tabularnewline
\bottomrule
\end{longtable}

@@

\begin{longtable}[c]{@{}lrr@{}}
\caption{Confusion matrix comparing chlorophyll \textit{a} predicted
exceedences (rows) versus real exceedances (columns) for the USEPA adult
drinking water advisory. \label{tab:adult_conmat_table}}\tabularnewline
\toprule
& FALSE & TRUE\tabularnewline
\midrule
\endfirsthead
\toprule
& FALSE & TRUE\tabularnewline
\midrule
\endhead
FALSE & 887 & 57\tabularnewline
TRUE & 50 & 34\tabularnewline
\bottomrule
\end{longtable}

@@

\begin{longtable}[c]{@{}lrr@{}}
\caption{Confusion matrix comparing chlorophyll \textit{a} predicted
exceedences (rows) versus real exceedances (columns) for the WHO
recreational water advisory.
\label{tab:who_rec_conmat_table}}\tabularnewline
\toprule
& FALSE & TRUE\tabularnewline
\midrule
\endfirsthead
\toprule
& FALSE & TRUE\tabularnewline
\midrule
\endhead
FALSE & 903 & 48\tabularnewline
TRUE & 50 & 27\tabularnewline
\bottomrule
\end{longtable}

@@

Heisler, J., P. M. Glibert, J. M. Burkholder, D. M. Anderson, W.
Cochlan, W. C. Dennison, Q. Dortch, C. J. Gobler, C. A. Heil, E.
Humphries, and others. 2008. Eutrophication and harmful algal blooms: A
scientific consensus. Harmful algae 8:3--13.
